---
Title: "Human Activity Recognition - Practical Machine Learning"
Author: "Saswat David"
Date: "June 21, 2020"
---
# Human Activity Recognition - Practical Machine Learning

## Introduction

Predicting quality of barbell lifts - classified from A to E, using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. Done as the course project of Practical Machine Learning course offered by John Hopkins University on Coursera.

## Import Libraries & Data

Requred packages are imported and workspace initialized. Seed is set to ensure reproducable results.

```{r}
#Removing pre-existing variables from workspace
rm(list = ls())

#Set Seed
set.seed(42)

#Initializing Libraries 
library(caret)
library(rattle)
library(rpart)
library(rpart.plot)
library(randomForest)
library(repmis)
```

Data is imported directly from the provided url.

```{r}
#Importing Data
trainurl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
validurl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

train <- source_data(trainurl, na.strings = c("NA", "#DIV/0!", ""), header = TRUE)
valid <- source_data(validurl, na.strings = c("NA", "#DIV/0!", ""), header = TRUE)
```

## Data Treatment

Columns with more than 80% NAs are removed

```{r}
#Remove Columns with > 80% NAs
train <- train[, colSums(is.na(train)) < 0.8*nrow(train)]
valid <- valid[, colSums(is.na(valid)) < 0.8*nrow(valid)]
```

Columns that won't help predict the target variable are removed.

```{r}
#Remove Redundant Columns
train <- train[, -c(1:7)]
valid <- valid[, -c(1:7)]
```

## Train - Test Split

Data is split with 80% of the records (15,699) in the training data and the remaining 20% (3,923) in the test dataset.

```{r}
#Train - Test Split
inTrain <- createDataPartition(train$classe, p = 0.8, list = FALS	E)
test <- train[-inTrain,]
train <- train[inTrain,]
```

## Model

Random Forest model is created with 10-Fold Cross-Validation.

```{r}
#Resampling Method
control <- trainControl("cv")

#Build Model
model <- train(classe ~ ., data = train, method = "rf", trControl = trainControl("cv"))

#View Model Details
print(model, digits = 4)
```

## Predict

Using created model, the 20% test dataset and 20 observation unknown dataset are predicted.

```{r}
#Predict Test & Unknown Dataset
predicted_t <- predict(model, test)
predicted_v <- predict(model, valid)

#20 Observation Unknown Dataset
prediction <- cbind(problem_id = valid$problem_id, prediction = as.character(predicted_v))
```

## Evaluate Model

The model is evaluated by looking at it's confusion matrix and accuracy numbers. The model has an accuracy of 99.4% on the test dataset.

```{r}
#Confusion Matrix
confusionMatrix(as.factor(test$classe), predicted_t)
```